\section{Conclusion}

\subsection{Conclusion}
This study has successfully developed and evaluated a real-time traffic signal violation detection system for the Simpang Wirosaban intersection. By integrating YOLO11 for object detection and the SORT algorithm for multi-object tracking, the system effectively monitors vehicle trajectories within a defined region of interest. The implementation of a dedicated three-stage image enhancement pipeline, comprising gamma correction, CLAHE, and unsharp masking, has proven effective in overcoming the challenges of nighttime illumination degradation, thereby increasing the system's overall detection robustness. Additionally, the decoupled HSV-based traffic light classifier ensures accurate state recognition, while the rule-based logic provides a reliable mechanism for identifying red light violations. The experimental results confirm that repurposing existing CCTV infrastructure with advanced computer vision techniques is a viable, cost effective solution for automated traffic enforcement.
\figViolationEx

\subsection{Future Work}
The following directions are identified for improving the system in subsequent research:
\begin{itemize}
    \item Training a custom vehicle detection model using deep learning on a domain-specific dataset collected from Indonesian intersections, to improve robustness against local vehicle types, road conditions, and camera perspectives. 
    \item Extending the system to support dynamic camera configurations, enabling deployment on pan-tilt-zoom (PTZ).
    \item Incorporating adverse weather robustness, particularly for rain conditions which introduce visual noise, lens distortion, and reduced contrast.
\end{itemize}