\section{Literature Review}

\subsection{Related Work}

Ong et al. \cite{ong2024} proposed a cost-effective traffic violation detection system using CCTV footage that targets two common violations: red light running and double-line overtaking. The system integrates YOLO for vehicle and traffic light detection, combined with a geometric approach using point-to-line distance to determine whether a vehicle has crossed a virtual boundary. Traffic light state is recognized directly from YOLO detections rather than a separate color-based classifier. The study demonstrated that vision-based approaches using existing CCTV infrastructure are viable alternatives to expensive sensor-based systems, achieving reliable detection across varied road conditions.

Tomas et al. \cite{tomas2024} developed a prototype system for detecting red light and wrong-way violations at intersections by combining YOLOv5 with Faster R-CNN. The system operates as an extension for existing CCTV installations, using object detection to identify vehicles and determine their relationship to the stop line and traffic signal state. Evaluated on intersection footage, the system achieved 100\% accuracy in detecting red light violations and 90\% accuracy for wrong-way violations. The study reinforces that a two-stage detection pipeline combining a fast single-stage detector with a more precise region-based network can improve overall system reliability for multi-class violation detection.

Mohan et al. \cite{mohan2023} presented a YOLOv8-based traffic violation detection system that integrates the SORT algorithm for multi-object tracking. The system assigns unique vehicle identities per frame, enabling consistent monitoring of vehicle trajectories across the detection zone. Upon identifying a red light violation, the system generates visual alerts and stores vehicle IDs for subsequent identification. The study highlights that pairing a modern single-stage detector with a lightweight tracker such as SORT provides a computationally practical framework for real-time enforcement, and identifies detection quality as the primary factor influencing overall tracking and violation classification performance.

The present study differs from the above works in three respects. First, YOLO11 is employed in place of earlier YOLO variants, offering higher mAP with fewer parameters. Second, a dedicated HSV-based traffic light classifier with temporal smoothing is implemented separately from the vehicle detector, improving robustness under nighttime illumination. Third, a three-stage image enhancement pipeline comprising gamma correction, CLAHE, and unsharp masking is applied as preprocessing to address the specific low-light conditions of the Simpang Wirosaban intersection.

\subsection{HSV (Hue, Saturation, Value) Color Model}
In the field of computer vision, selecting an appropriate color space is critical for the accuracy of object detection under varying environmental conditions. The HSV (Hue, Saturation, Value) color model is a cylindrical representation of colors that more closely aligns with human visual perception compared to the standard RGB model. While RGB defines colors as a combination of primary light sources, it is highly sensitive to changes in brightness, shadows, and external lighting.The HSV model decouples color information from intensity, using three distinct components:
\begin{itemize}
    \item Hue ($H$): Represents the dominant wavelength or the actual "color" (e.g., Red, Yellow, Green) expressed in degrees from $0^{\circ}$ to $360^{\circ}$.
    \item Saturation ($S$): Indicates the vibrancy or purity of the color, ranging from $0$ to $100\%$.
    \item Value ($V$): Represents the brightness or luminance of the color.
\end{itemize}

\subsection{You Only Look Once 11 (YOLO11)}

YOLO11 is a single-stage object detection model released by Ultralytics in September 2024 at the YOLO Vision 2024 event \cite{ultralytics2024}. It follows the conventional backbone–neck–head architecture common to the YOLO family, with three principal architectural innovations that distinguish it from its predecessors.

The backbone employs C3k2 blocks (Cross Stage Partial with kernel size 2) as a replacement for the C2f blocks used in YOLOv8. C3k2 implements the CSP bottleneck using two convolutions with smaller kernels, reducing parameter count while preserving gradient flow across stages \cite{khanam2024yolo11}. Multi-scale feature aggregation is handled by the Spatial Pyramid Pooling Fast (SPPF) module in the neck, which applies sequential max-pooling operations of fixed kernel size and concatenates the results, enabling the model to capture context at multiple receptive field scales without the computational overhead of the original SPP formulation. Following SPPF, the C2PSA module (Convolutional block with Parallel Spatial Attention) incorporates a self-attention mechanism that allows the model to selectively weight spatial regions of the feature map, improving sensitivity to small and partially occluded objects \cite{khanam2024yolo11}. The detection head retains the decoupled structure introduced in YOLOv8, maintaining separate branches for classification and bounding box regression.

\figYOLOArchitecture

Compared to YOLOv8m, YOLO11m achieves higher mean Average Precision (mAP) on the COCO benchmark while using 22\% fewer parameters \cite{ultralytics2024}, making it well suited for real-time deployment on GPU hardware. In this study, the medium variant (YOLO11m) is used with inference restricted to five COCO vehicle classes: bicycle, car, motorcycle, bus, and truck.

\subsection{Simple Online and Realtime Tracking (SORT)}

SORT is a multi-object tracking algorithm proposed by Bewley et al. \cite{bewley2016} that follows the tracking-by-detection paradigm, where an external detector supplies bounding boxes each frame and the tracker is responsible solely for associating detections across time. Despite relying exclusively on bounding box geometry without any appearance features, SORT achieves tracking accuracy comparable to far more complex methods while operating at over 260 Hz \cite{bewley2016}.

Each tracked object is represented by a \texttt{KalmanBoxTracker} with a 7-dimensional state vector:

\begin{equation}
\mathbf{x} = [c_x,\ c_y,\ s,\ r,\ \dot{c}_x,\ \dot{c}_y,\ \dot{s}]^T
\end{equation}

\noindent where $c_x$ and $c_y$ are the centroid coordinates of the bounding box, $s$ is the box area, $r$ is the aspect ratio, and $\dot{c}_x$, $\dot{c}_y$, $\dot{s}$ are their respective first-order velocities. The aspect ratio $r$ is assumed constant. The Kalman filter employs a constant velocity linear motion model as the state transition, with the observation vector consisting of $[c_x, c_y, s, r]^T$ only. At each frame, every active tracker advances its state estimate through the Kalman prediction step before any new detections are considered.

Association between predicted tracker states and incoming detections is formulated as a linear assignment problem. An IoU cost matrix is constructed between all predicted bounding boxes and all current detections, and the optimal one-to-one assignment is solved using the Hungarian algorithm \cite{bewley2016}. A detection–tracker pair is accepted only if their IoU exceeds a minimum threshold; pairs below this threshold are treated as unmatched. Matched trackers update their Kalman states with the observed measurement. Unmatched detections initialize new trackers, and trackers that remain unmatched beyond a maximum age threshold are removed. A tracker is only reported as a confirmed track once it has been matched for a minimum number of consecutive frames, suppressing spurious short-lived detections.
