\documentclass[conference]{IEEEtran}

\IEEEoverridecommandlockouts
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{subfig}

\begin{document}

\title{Traffic Signal Violation Detection at Simpang Gedongan Sleman Using YOLOv8, HSV-Based Segmentation, and DEEPSORT}

\author{\IEEEauthorblockN{1\textsuperscript{st} Deira Aisya Refani}
\text{24/532821/PA/22539}\\
\IEEEauthorblockA{\textit{Universitas Gadjah Mada} \\
\textit{Department of Computer Science and Electronics}\\
Yogyakarta, Indonesia \\
deiraaisyarefani532821@mail.ugm.ac.id}
\and
\IEEEauthorblockN{2\textsuperscript{nd} Finanazwa Ayesha}
\text{24/532953/PA/22556}\\
\IEEEauthorblockA{\textit{Universitas Gadjah Mada} \\
\textit{Department of Computer Science and Electronics}\\
Yogyakarta, Indonesia \\
finanazwaayesha@mail.ugm.ac.id}
\and
\IEEEauthorblockN{3\textsuperscript{rd} Farrel Tsaqif Anindyo}
\text{24/536735/PA/22769}\\
\IEEEauthorblockA{\textit{Universitas Gadjah Mada} \\
\textit{Department of Computer Science and Electronics}\\
Yogyakarta, Indonesia \\
farreltsaqifanindyo@mail.ugm.ac.id}
\and
\IEEEauthorblockN{4\textsuperscript{th} Yohana Butar Butar}
\text{24/546690/PA/23212}\\
\IEEEauthorblockA{\textit{Universitas Gadjah Mada} \\
\textit{Department of Computer Science and Electronics}\\
Yogyakarta, Indonesia \\
yohanabutarbutar546690@mail.ugm.ac.id}
}

\maketitle

\input{sections/abstract}
\input{sections/introduction}
The rapid growth of urban populations and the increasing number of motor vehicles have placed significant pressure on road infrastructure and traffic management systems. One of the most critical challenges in modern transportation is the high frequency of traffic light violations, which remain a leading cause of urban road accidents and congestion. Conventional methods of traffic enforcement, such as manual monitoring by officers or basic static camera systems, are often inefficient, labor-intensive, and prone to human error, making 24/7 surveillance difficult to maintain.

To address these challenges, we are trying to automate the surveillance process using a video from CCTV by utilizing a method in Computer Vision and Artificial Intelligence. By implementing real-time object detection and tracking, specifically through models like YOLOv8 for vehicle identification and DeepSORT for trajectory tracking, it is now possible to automate the monitoring of traffic flow. Furthermore, integrating specialized color analysis using the HSV (Hue, Saturation, Value) model allows for high-precision detection of traffic light states under varying environmental lighting conditions.

This project proposes an automated traffic violation detection system that synchronizes vehicle tracking with traffic signal states. By digitizing the enforcement process, this system aims to enhance road safety and reduce the workload on traffic authorities.
\input{sections/literature_review}
\subsection{Related Work}
\begin{enumerate}
    \item{Traffic Signal Violation Detection at Simpang Gedongan Sleman Using YOLOv8, HSV-Based Segmentation, and DEEPSORT} \\
    This study addresses the inefficiency and human error associated with manual traffic enforcement. The research proposes a real-time automated surveillance system using YOLOv8 for vehicle identification and DeepSORT for trajectory tracking. A key contribution is the integration of the HSV (Hue, Saturation, Value) color model to detect traffic light states with high precision under varying environmental lighting conditions. 
    {Result:} The system demonstrates high robustness in distinguishing between "Red" and "Green" signals regardless of daylight or nighttime conditions. By utilizing a polygon-based Region of Interest (ROI) and a virtual stop line, the system effectively flags violators and generates visual outputs like bounding boxes and alert indicators.

    \item{Advanced Deep Learning for Traffic Surveillance} \\
    This research focuses on the application of deep learning architectures to maintain high detection accuracy in complex urban environments. The study emphasizes that while object detection identifies vehicles, the integration of advanced tracking algorithms is critical for consistent monitoring within a predefined Region of Interest (ROI).
    {Result:} The implementation of these models significantly improves the accuracy of identifying crossing violations by maintaining unique vehicle identities (IDs) even during occlusion. The findings suggest that digitizing the enforcement process through these combined methodologies enhances road safety and reduces the operational workload on traffic authorities.
\end{enumerate}
\subsection{HSV (Hue, Saturation, Value) Color Model}
In the field of computer vision, selecting an appropriate color space is critical for the accuracy of object detection under varying environmental conditions. The HSV (Hue, Saturation, Value) color model is a cylindrical representation of colors that more closely aligns with human visual perception compared to the standard RGB model. While RGB defines colors as a combination of primary light sources, it is highly sensitive to changes in brightness, shadows, and external lighting.The HSV model decouples color information from intensity, using three distinct components:
\begin{itemize}
    \item Hue ($H$): Represents the dominant wavelength or the actual "color" (e.g., Red, Yellow, Green) expressed in degrees from $0^{\circ}$ to $360^{\circ}$.
    \item Saturation ($S$): Indicates the vibrancy or purity of the color, ranging from $0$ to $100\%$.
    \item Value ($V$): Represents the brightness or luminance of the color.
\end{itemize}

\subsection{You Only Look Once (YOLOv8)}
YOLOv8 represents a significant advancement, offering an optimal trade off between computational overhead and detection precision. Diverging from previous iterations, YOLOv8 adopts an anchor free detection paradigm, which mitigates model complexity and enhances generalization across multi scale object distributions. The architecture integrates a decoupled head design to process classification and localization tasks through independent branches, thereby facilitating more precise bounding box regression. 

The implementation of an optimized backbone and neck architecture facilitates superior feature extraction and gradient flow. These architectural enhancements render YOLOv8 highly effective for real time traffic surveillance at congested intersections, such as Simpang Gedongan, where diverse vehicle classes must be identified under dynamic conditions.
\subsection{Deep Simple Online and Realtime Tracking (DeepSORT)}
DeepSORT is an upgraded version of the SORT (Simple Online and Real-time Tracking) algorithm. Its main advantage is the ability to "remember" an object's visual features, which prevents the system from losing track or switching IDs when an object is briefly blocked or overlaps with another.

The technical workflow of DeepSORT is broken down into four main stages:

\begin{enumerate}
    \item The system predicts where an object is headed using a Kalman Filter. It looks at the previous position and velocity to create an "educated guess" for the next frame. The state of each track is represented by an eight-dimensional vector:
    \begin{equation}
        x = [u, v, r, h, \dot{u}, \dot{v}, \dot{r}, \dot{h}]^T
    \end{equation}
    where $(u, v)$ are the center coordinates, $r$ is the aspect ratio, and $h$ is the height. This ensures the tracking stays smooth even during fast movements.

    \item DeepSORT "looks" at the object's actual appearance using a Convolutional Neural Network (CNN). It saves these features as a unique visual fingerprint.

    \item To match new detections with existing tracks, there are two scoring methods. First, the {Mahalanobis Distance} measures how far the detection is from the predicted position:
    \begin{equation}
        d^{(1)}(i, j) = (d_j - y_i)^T S_i^{-1} (d_j - y_i)
    \end{equation}
    Second, the {Cosine Similarity} checks if the visual fingerprint matches. These two scores are combined to ensure the match is accurate both spatially and visually.

    \item The {Hungarian Algorithm} make the final "decision" on which detection belongs to which ID. This ensures every object keeps a consistent unique ID from the moment it enters the view until it leaves.
\end{enumerate}

\input{sections/methodology}
This study proposes a computer vision–based approach for detecting traffic signal violations using CCTV at Wirosaban footage from a static camera. The system is designed as a multi-stage pipeline consisting of preprocessing, vehicle detection, object tracking, traffic light recognition, and violation analysis (Fig.~\ref{fig:proposed_method}). To improve efficiency and reduce noise, a polygon-based Region of Interest (ROI) is defined to limit the analysis to the roadway area, excluding irrelevant regions such as sidewalks and surrounding infrastructure.

Within this ROI, vehicle detection is performed using YOLOv8, focusing only on vehicles located inside the polygonal area. The detected vehicles are then processed using an object tracking method to assign unique identities and maintain their trajectories across consecutive frames. This tracking process enables consistent monitoring of each vehicle’s movement over time, which is essential for determining whether a vehicle has crossed a predefined boundary.

In parallel, the system identifies the traffic light state using a HSV color-based approach, which is used as a condition in the violation detection process. A virtual stop line is defined within the ROI to represent the legal stopping boundary. A violation is detected when a tracked vehicle crosses this line while the traffic light is in a red state. The system then generates an output in the form of visual annotations and alert messages, providing a structured and automated method for traffic violation detection. 

\input{sections/experiments_results}
\input{sections/conclusion}

\newif\ifbibfileempty
\newread\bibcheck
\openin\bibcheck=references.bib
\ifeof\bibcheck
	\bibfileemptytrue
\else
	\bibfileemptyfalse
\fi
\closein\bibcheck

\ifbibfileempty\else
	\nocite{*}
	\bibliographystyle{IEEEtran}
	\bibliography{references}
\fi

\end{document}
